### 2025-04-10 Live_Scripts/R_version_Rubisco_activation/Scripts/Temperature-response-of-Rubisco-activation-in-rice.Rmd
Add files via upload

---
title: "Temperature response of Rubisco activation state in rice"
author: "Supreeta Vijayakumar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(readxl)
library(ggplot2)
```

## Data Extraction

To modify the regulation of Rubisco in the e-Photosynthesis model, we need to fit a model to be able to predict the response of activation state in rice to changes in temperature.

\
There is published data in Fig 4A of [Makino and Sage (2007)](https://academic.oup.com/pcp/article/48/10/1472/1938331) which plots the change in activation state across temperatures for rice:\
\
![](Data/Makino_Sage_Fig_4-01.png)

To be able to extract this data, the points can be estimated manually using measuring tools or software such as [GraphReader](https://www.graphreader.com/), [PlotDigitizer](https://plotdigitizer.com/app) or [WebPlotDigitizer](https://automeris.io/).\
\
Define the data points:

```{r load data}
## Use manual estimates
# x <-c(16,20,25,32,37,41) # Temperature
# y <-c(85,85,80,73,69,48) # Activation state (%)

# Load data from .csv file compiled from automated tools

# These points are extracted using Plot Digitizer
#plot_data1 <- read.csv("Data/Fig_4A-plot-data.csv")
#x1 <- plot_data$x # Temperature
#y1 <- plot_data$y # Activation state (%)

# These points are extracted using Web Plot Digitizer
plot_data <- read.csv("Data/Fig_4A_plot_data_2.csv", header = FALSE)

# Convert activation state from percentage to decimal
plot_data$V2<- plot_data$V2/100

# Specify variables
x <- plot_data$V1 # Temperature
y <- plot_data$V2 # Activation state (%)

print(plot_data)
```

Plot the points, setting y-axis breaks at 20% intervals:

```{r plot data}
ggplot(plot_data, 
       aes(x, y)) +
       geom_line(color = "black", 
                 linewidth = 1) +
       geom_point(color = "black", 
                  fill = "black", 
                  shape = 21, 
                  size = 2) +
       scale_x_continuous(limits = c(10, 45)) +
       scale_y_continuous(limits = c(0, 1.2), 
                          breaks = seq(0.2, 1.2, by = 0.2)) +
       labs(x = "Leaf temperature (°C)", 
            y = "Rubisco activation (%)") +
       theme_minimal()
```

## Polynomial curve fitting

Try fitting the data to quadratic, cubic and quartic models:

```{r fit polynomials}
poly_model2 <- lm(y ~ poly(x, 
                           2, 
                           raw = TRUE),
                      data = plot_data)
poly_model3 <- lm(y ~ poly(x, 
                           3, 
                           raw = TRUE),
                      data = plot_data)
poly_model4 <- lm(y ~ poly(x, 
                           4, 
                           raw = TRUE),
                      data = plot_data)
```

## Evaluation of fitted models

Check the summary statistics for the three fits:

```{r summarise polynomial fit stats}
summary_quad <- summary(poly_model2)
summary_quad

summary_cubic <- summary(poly_model3)
summary_cubic

summary_quartic <- summary(poly_model4)
summary_quartic
```

The coefficients for the 4th-degree polynomial indicate a very steep model, which could cause values to escalate or drop rapidly, especially near the boundaries or beyond the range of the data. So this is not likely to be an appropriate model, especially for fitting only 6 data points.\
\
When comparing fit summaries of the quadratic and cubic models, the cubic model explains a higher percentage of the variability (96.16% vs. 93.73%).

The cubic model has a slightly lower RSE and slightly higher adjusted $R^{2}$ value, so it fits slightly better to the data than the quadratic as it includes an extra term - adding to complexity. On the other hand, the quadratic model doesn't explain as much variance, but it is penalised less for including fewer predictors. It is also better than the cubic model in terms of statistical significance (P-val sig \**.*\* for intercept and quadratic term in the quadratic model = significant at 10% level vs. none of the terms being significant for the cubic model).

Extract the fit coefficients for the quadratic and cubic models:

```{r extract polynomial coefficients}
coef_quad <- coef(poly_model2)
coef_quad
coef_cubic <- coef(poly_model3)
coef_cubic
```

Extract the residuals for the quadratic and cubic models:

```{r extract residuals}
residuals_quad <- residuals(poly_model2)
residuals_cubic <- residuals(poly_model3)
```

Save the polynomial functions including the model coefficients for the quadratic and cubic models:

```{r save polynomial functions}
quad_func <- function(x) coef_quad[1] + 
                          coef_quad[2] * x + 
                          coef_quad[3] * x^2
cubic_func <- function(x) coef_cubic[1] + 
                          coef_cubic[2] * x + 
                          coef_cubic[3] * x^2 + 
                          coef_cubic[4] * x^3
```

Obtain formulae for the polynomial functions:

```{r polynomial formulae}
quad_formula <- bquote(Quadratic: y == 
                       .(round(coef_quad[3], 4)) * x^2 + 
                       .(round(coef_quad[2], 3)) * x + 
                       .(round(coef_quad[1], 2)))

cubic_formula <- bquote(Cubic: y == 
                        .(round(coef_cubic[4], 5)) * x^3 + 
                        .(round(coef_cubic[3], 4)) * x^2 + 
                        .(round(coef_cubic[2], 3)) * x + 
                        .(round(coef_cubic[1], 2)))
quad_formula
cubic_formula
```

The `bquote` function ensures creation of the formulae with LaTeX formatting.

`.()` is used to evaluate the expressions and insert the coefficients dynamically into the formulae.

Plot the polynomial models on top of the data points using stat_function:

```{r plot polynomials}
ggplot(plot_data, 
       aes(x, y)) +
  geom_line(color = "black", 
            linewidth = 1) +
  geom_point(color = "black", 
             fill = "black", 
             shape = 21, 
             size = 2) +
  stat_function(fun = quad_func, 
                aes(color = "Quadratic"), 
                linewidth = 1, 
                linetype = "dashed",
                xlim = c(0, 45)) +  
  stat_function(fun = cubic_func, 
                aes(color = "Cubic"), 
                linewidth = 1, 
                linetype = "dashed",
                xlim = c(0, 45)) +  
  scale_x_continuous(limits = c(10, 45)) +
  scale_y_continuous(limits = c(0, 1.2), 
                     breaks = seq(0.2, 1.2, by = 0.2)) +
  labs(x = "Leaf temperature (°C)", 
       y = "Rubisco activation (%)",
       color = "Polynomial Fit") +
  theme_minimal() +
  scale_color_manual(
    values = c("Quadratic" = "red", 
               "Cubic" = "blue"),
    labels = c(quad_formula, 
               cubic_formula))
```

Obtain predictions using the polynomial models:

```{r predict with polynomial models}
y_pred_quad <- predict(poly_model2)
y_pred_cubic <- predict(poly_model3)
```

Plot the residuals against temperature for the quadratic model:

```{r plot quad residuals vs. temp}
ggplot(plot_data, 
       aes(x = x)) +
       geom_point(aes
                  (y = residuals_quad), 
                  color = "red") +
       geom_hline(yintercept = 0,
                  linetype = "dashed") +
       ggtitle("Residuals of Quadratic Model vs. Temperature") +
       xlab("Leaf temperature (°C)") +
       ylab("Residuals") +
       theme_minimal()
```

Plot the residuals against temperature for the cubic model:

```{r plot cubic residuals vs. temp}
ggplot(plot_data, 
       aes(x = x)) +
       geom_point(aes
                  (y = residuals_cubic), 
                  color = "blue") +
       geom_hline(yintercept = 0,
                  linetype = "dashed") +
       ggtitle("Residuals of Cubic Model vs. Temperature") +
       xlab("Leaf temperature (°C)") +
       ylab("Residuals") +
       theme_minimal()
```

Plot the residuals against fitted values for the quadratic model:

```{r plot quad residuals vs. fitted}
ggplot(plot_data, 
       aes(x = y_pred_quad)) +
       geom_point(aes
                  (y = residuals_quad), 
                  color = "red") +
       geom_hline(yintercept = 0,
                  linetype = "dashed") +
       ggtitle("Residuals of Quadratic Model vs. Fitted Values") +
       xlab("Fitted Values") +
       ylab("Residuals") +
       theme_minimal()
```

Plot the residuals against fitted values for the cubic model:

```{r plot cubic residuals vs. fitted}
ggplot(plot_data, 
       aes(x = y_pred_cubic)) +
       geom_point(aes
                  (y = residuals_cubic), 
                  color = "blue") +
       geom_hline(yintercept = 0,
                  linetype = "dashed") +
       ggtitle("Residuals of Cubic Model vs. Fitted Values") +
       xlab("Fitted Values") +
       ylab("Residuals") +
       theme_minimal()
```

There is no observable pattern in the scatter of residuals and they are close to the RSE values, which suggests the models are appropriate and have captured the underlying relationship between leaf temperature and activation state - even if some of the residual values are a bit higher, they are all \<10.

Compute MSE, RMSE and AIC for the models:

```{r compute mse and aic}
mse_quad <- mean((y - y_pred_quad)^2)
mse_quad
mse_cubic <- mean((y - y_pred_cubic)^2)
mse_cubic
rmse_quad <- sqrt(mean((y - y_pred_quad)^2))
rmse_quad
rmse_cubic <- sqrt(mean((y - y_pred_cubic)^2))
rmse_cubic
aic_quad <- AIC(poly_model2)
aic_quad
aic_cubic <- AIC(poly_model3)
aic_cubic
```

The cubic has a lower MSE and AIC but there is not that much difference between the two AIC values.