### 2025-04-10 Live_Scripts/R_version_Vmax_temp/Scripts/Vmax_temp.Rmd
Add files via upload

---
title: "Temperature-dependent adjustment of $V_{c}$ and $V_{o}$"
author: "Supreeta Vijayakumar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load any libraries here
library(ggplot2)
library(minpack.lm)
```

# Bernacchi temperature functions

Currently the e-Photosynthesis model sets PrV111 = PsV1 \* 0.24, according to the known value for tobacco at 25°C (Whitney et al., 1999).\

However, this ratio varies across different crop species and temperatures.\
Seeing as there is no existing temperature functions for rice, we will use the equations by Bernacchi (2001) and values from Makino (1988) to compare the temperature responses of the parameters.

Specify constants:

```{r}
T25 <- 25 + 273.15                 # Reference temperature in K 
Tp <- 28.9310407291759 + 273.15    # Average measurement temperature in K
R <- 0.008314                     # Ideal gas constant in kJ K-1 mol-1
c_Vc <- 26.35                       # Scaling constant, Vcmax             (N. tabacum L. cv. W38, Bernacchi et al 2001)
dHa_Vc <- 65.33                    # Activation energy, Vcmax, kJ mol -1 (N. tabacum L. cv. W38, Bernacchi et al 2001)
c_Vo <- 22.98                       # Scaling constant, Vomax             (N. tabacum L. cv. W38, Bernacchi et al 2001)
dHa_Vo <-60.11                     # Activation energy, Vomax, kJ mol -1 (N. tabacum L. cv. W38, Bernacchi et al 2001)
```

Use the Arrhenius equation to calculate the $\frac {V_{o}} {V_c}$ ratio at 25°C for N.
tabacum (c and dHa from Bernacchi et al., 2001):\

$$Parameter = exp (c- \frac{\Delta H_{a}} {R \times T_{K}})$$\

```{r}
Bernacchi_Vc_25 <- exp(c_Vc - dHa_Vc / (R * T25))
Bernacchi_Vo_25 <- exp(c_Vo - dHa_Vo / (R * T25))
Bernacchi_PrPs_ratio_25 <- Bernacchi_Vo_25 / Bernacchi_Vc_25
print(Bernacchi_PrPs_ratio_25)
```

Calculate $\frac {V_{o}} {V_c}$ ratios across a range of temperatures (10-40°C) :

```{r}
Temp_Range <- data.frame(Temp_C = 10:40)                       # Temperatures in °C
Temp_Range$Temp_K <- Temp_Range$Temp_C + 273.15                # Temperatures in Kelvin
Temp_Range$Vc <- exp(c_Vc - dHa_Vc / (R * Temp_Range$Temp_K))  # Vc values
Temp_Range$Vo <- exp(c_Vo - dHa_Vo / (R * Temp_Range$Temp_K))  # Vo values
Temp_Range$Vo_Vc_ratio <- Temp_Range$Vo / Temp_Range$Vc        # Vo/Vc ratios
```

Plot $\frac {V_{o}} {V_c}$ data:

```{r}
ggplot(Temp_Range, aes(x = Temp_C, y = Vo_Vc_ratio)) +
                   geom_line(color = "blue") +
                   geom_point(color = "blue", size = 2) +
                   labs(x = expression("Temperature ("*degree*C*")"), y = "Vo/Vc") +
                   theme_minimal()
ggsave("Outputs/Temp_vs_VcVo.png")
```

Plot $V_{o}$:

```{r}
ggplot(Temp_Range, aes(x = Temp_C, y = Vo)) +
                   geom_line(color = "green") +
                   geom_point(color = "green", size = 2) +
                   labs(x = expression("Temperature ("*degree*C*")"), y = "Vo") +
                   theme_minimal()
ggsave("Outputs/Temp_vs_Vo.png")
```

Plot $V_{c}$:

```{r}
ggplot(Temp_Range, aes(x = Temp_C, y = Vc)) +
                   geom_line(color = "red") +
                   geom_point(color = "red", size = 2) +
                   labs(x = expression("Temperature ("*degree*C*")"), y = "Vc") +
                   theme_minimal()
ggsave("Outputs/Temp_vs_Vc.png")
```

Calculate Bernacchi values at measurement temperature:

```{r}
Bernacchi_Vc_Tp <- exp(c_Vc - dHa_Vc / (R * Tp))
Bernacchi_Vo_Tp <- exp(c_Vo - dHa_Vo / (R * Tp))
Bernacchi_PrPs_ratio_Tp <- Bernacchi_Vo_Tp / Bernacchi_Vc_Tp
print(Bernacchi_PrPs_ratio_Tp)
```

Calculate Makino values at 25°C:

```{r}
Makino_Vc_25 <- 1.77                                # Carboxylase activity (µmol (mg enzyme)⁻¹ min⁻¹)
Makino_Vo_25 <- 0.58                                # Oxygenase activity (µmol (mg enzyme)⁻¹ min⁻¹)
Makino_PrPs_ratio_25 <- Makino_Vo_25 / Makino_Vc_25 # Vo/Vc ratio
print(Makino_PrPs_ratio_25)
```

## Adjusting the value $V_{cmax}$ at 25°C to measurement temperature $T{p}$

We can use one of two approaches to calculate the Makino value of $\frac {V_{o}} {V_c}$ at Tp:

1)  Use the ratios of Makino and Bernacchi at 25°C and multiply by Bernacchi parameter estimate at a given temperature

2)  Fit an Arrhenius equation to a non-linear model to find the Bernacchi temperature response of (plotted above)

Use linear scaling of Makino to Bernacchi ratios:

```{r}
Makino_PrPs_ratio_Tp <- Makino_PrPs_ratio_25*Bernacchi_PrPs_ratio_Tp/Bernacchi_PrPs_ratio_25
print(Makino_PrPs_ratio_Tp)
```

## Non-linear fits for $\frac {V_{o}} {V_c}$

Define predictor and response variables:

```{r}
T <- Temp_Range$Temp_K      # Temperature in K
y <- Temp_Range$Vo_Vc_ratio # Vo/Vc ratio
```

So far, we have defined the Arrhenius equation as:

$$Parameter = exp (c- \frac{\Delta H_{a}} {R \times T_{K}})$$\

$c$ is meant to scale the amplitude of the function whereas $\Delta H_{a}$ represents the temperature dependence relationship.

Using the term $c−\Delta H_{a}$ combined these two effects into one, making them harder to interpret individually.
Without $c$, the model assumes that $\Delta H_{a}$ depends solely on temperature and ignores other potential variations.

This means the equation can also be rewritten as

$$y = c \times exp(-\frac{\Delta H_{a}} {R \times T_{K}})$$\

to ensure that the exponent has a negative sign - previously $exp(\frac{c- \Delta H_{a}} {R \times T_{K}})$ was causing errors since the exponent was positive.

Define an exponential model with non-linear parameters using this form of the Arrhenius equation.
Start by creating a fit type for a custom nonlinear model using nls():

```{r}
#nls_model <- nls(y ~ c_VoVc * exp(-dHa_VoVc / (R * T)), data = Temp_Range, start = list(c_VoVc = 0.3, dHa_VoVc = 60))
```

This is producing errors so use nlsLM() - which uses the Levenberg-Marquardt algorithm instead of the Gauss-Newton used by nls() for nonlinear least-squares estimation - for a more robust fitting process:

```{r}
nls_model <- nlsLM(Vo_Vc_ratio ~ c_VoVc * exp(-dHa_VoVc / (R * T)),
                                 data = Temp_Range,
                                 start = list(c_VoVc = 0, 
                                              dHa_VoVc = 2))
```

View the plot summary:

```{r}
summary(nls_model)
```

nlsLM() converged successfully so we can add the fitted values and new parameter estimates for c and dHa to Temp_Range dataframe:

```{r}
Temp_Range$fitted <- predict(nls_model)
```

Plot the non-linear fit of the model to the data:

```{r}
ggplot(Temp_Range, aes(x = Temp_K, y = Vo_Vc_ratio)) +
                   geom_point() +
                   geom_line(aes(y = fitted), color = "red") +
                   labs(x = "Temperature (K)", 
                        y = "Vo/Vc Ratio", 
                        title = "Non-linear Fit for Vo/Vc")
ggsave("Outputs/Temp_vs_VoVc_nlsLM.png")
```

Calculate the residuals by subtracting the fitted values from the observed responses:

```{r}
Temp_Range$nls_residuals <- Temp_Range$Vo_Vc_ratio - Temp_Range$fitted
```

Plot residuals against temperature:

```{r}
ggplot(Temp_Range, aes(x = Temp_K, y = nls_residuals)) +
                       geom_point() +
                       geom_hline(yintercept = 0, 
                                  linetype = "dashed", 
                                  color = "red") +
                       scale_y_continuous(limits = c(-8e-16, +8e-16)) +
                       labs(x = "Temperature (K)", 
                            y = "Residuals", 
                            title = "Non-Linear Residuals vs. Temperature")
ggsave("Outputs/nlsLM_residuals_temp.png")
```

Plot residuals against fitted values:

```{r}
ggplot(Temp_Range, aes(x = fitted, y = nls_residuals)) +
                       geom_point() +
                       geom_hline(yintercept = 0, 
                                  linetype = "dashed", 
                                  color = "red") +
                       scale_y_continuous(limits = c(-8e-16, +8e-16)) +
                       labs(x = "Fitted Values", 
                            y = "Residuals", 
                            title = "Non-Linear Residuals vs. Fitted Values")
ggsave("Outputs/nlsLM_residuals_fitted.png")
```

Random scatter around zero indicates we have a good fit of model to data.
The values of the residuals are also very small ($<8 \times 10^{-16}$), indicating the model accounts for all variations in the data and there are no unexplained residuals left.

Calculate R-squared:

```{r}
SST <- sum((Temp_Range$Vo_Vc_ratio - mean(Temp_Range$Vo_Vc_ratio))^2)     # Total sum of squares
SSR <- sum(Temp_Range$nls_residuals^2)                                    # Residual sum of squares
R_squared <- 1 - (SSR / SST)
print(R_squared)
```

The R-squared is exactly 1, which indicates all the data points lie on the fitted curve perfectly.
This is expected because the data points being fit are predictions from fitted models for $V_{c}$ and $V_{o}$ rather than actual data, so they have no associated measurement error.

However, R-squared is not always recommended for non-linear regression, because non-linear models do not usually have an intercept term and the total sum of squares is not necessarily equal to the sum of the regression sum of squares and the residual sum of squares.

Alternatively, we can calculate the mean squared error (MSE) or Akaike Information Criterion (AIC) instead:
```{r}
squared_residuals <- Temp_Range$nls_residuals^2

mse <- mean(squared_residuals)
print(mse)

nls_AIC <- AIC(nls_model)
print(nls_AIC)
```

The mse value is very low (\<$1\times10^{-31}$), indicating that the model’s predictions are very close to the data points.
Likewise, lower AIC values also indicate a better model with a good balance between fit and complexity.

Fit a linear model for comparison:
```{r}
linear_model <- lm(Vo_Vc_ratio ~ Temp_K, data = Temp_Range)
summary(linear_model)

linear_AIC <- AIC(linear_model)
print(linear_AIC)

Temp_Range$lin_fitted <- predict(linear_model)
```

Plot the linear fit of the model to the data:
```{r}
ggplot(Temp_Range, aes(x = Temp_K, y = Vo_Vc_ratio)) +
                   geom_point() +
                   geom_line(aes(y = lin_fitted), color = "red") +
                   labs(x = "Temperature (K)", 
                        y = "Vo/Vc Ratio", 
                        title = "Linear Fit for Vo/Vc")
ggsave("Outputs/Temp_vs_VoVc_linear.png")
```

The simpler linear model seems to perform almost as well as the more complex non-linear model, but does not predict correctly across the entire range. We can see that its AIC value is much higher, indicating a poorer predictive model.
However, in some cases it would be have been advantageous to simplify the model e.g. if we wanted to avoid overfitting.

Extract the nlsLM derived parameter coefficients:

```{r}
coefficients <- coef(nls_model)
c_VoVc <- as.numeric(coefficients["c_VoVc"])
dHa_VoVc <- as.numeric(coefficients["dHa_VoVc"])
```

Work out the conversion ratio between Bernacchi and Makino Vo/Vc ratios at 25°C:

```{r}
conv_ratio <- Makino_PrPs_ratio_25/Bernacchi_PrPs_ratio_25
```

Predict Vo/Vc value at Tp using the estimated parameters:

```{r}
Bernacchi_PrPs_ratio_Tp2 <- (c_VoVc * exp(-dHa_VoVc / (R * Tp)))
print(Bernacchi_PrPs_ratio_Tp2)
```

Correct for the conversion to Makino:

```{r}
Makino_PrPs_ratio_Tp2 <- Bernacchi_PrPs_ratio_Tp2*conv_ratio
print(Makino_PrPs_ratio_Tp2)
```

Compare the two values obtained for Makino ratios at Tp and print TRUE if they are equal:

```{r}
are_equal <- Makino_PrPs_ratio_Tp == Makino_PrPs_ratio_Tp2
print(are_equal)  
```

Although are_equal is FALSE, the Tp2 value is essentially the same (upto 15 dp) as the Tp value calculated using Method 1) multiplying the ratios of Makino and Bernacchi at 25°C by the Bernacchi parameter estimate at Tp.

### Normalization fitting method

We can also compare the values by re-scaling the $\frac {V_{o}} {V_c}$ ratio from Bernacchi so that it has a value of 1 at 25°C and then fitting the Arrhenius function using these values.

Divide all $\frac {V_{o}} {V_c}$ values calculated at different temperatures by the value at 25°C:

```{r}
normalized_ratios <- Temp_Range$Vo_Vc_ratio/Bernacchi_PrPs_ratio_25;
```

Obtain Arrhenius constants using the re-scaled values:

```{r}
nls_norm_model <- nlsLM(normalized_ratios ~ c_VoVc * exp(-dHa_VoVc / (R * T)),
                                            data = Temp_Range,
                                            start = list(c_VoVc = 0, 
                                                         dHa_VoVc = 2))
summary(nls_norm_model)
```

Extract the coefficients from this second model:

```{r}
norm_coefficients <- coef(nls_norm_model)
c_norm <- as.numeric(norm_coefficients["c_VoVc"])
dHa_norm <- as.numeric(norm_coefficients["dHa_VoVc"])
```

Predict the Bernacchi ratio at Tp using the parameters obtained from the non-linear fit of normalized values:

```{r}
Bernacchi_PrPs_ratio_Tp3 <- (c_norm * exp(-dHa_norm / (R * Tp)))
print(Bernacchi_PrPs_ratio_Tp3)
```

This ratio is different from the other two ratios since we re-scaled the values and obtained a new parameter estimate for $c$.

Multiply the Bernacchi ratio at Tp by the Makino ratio at 25°C:

```{r}
Makino_PrPs_ratio_Tp3 <- Makino_PrPs_ratio_25*Bernacchi_PrPs_ratio_Tp3
print(Makino_PrPs_ratio_Tp3)
```

Compare the third value obtained for Makino ratios to the first two and print TRUE if equal:

```{r}
are_equal <- (Makino_PrPs_ratio_Tp == Makino_PrPs_ratio_Tp3) | 
             (Makino_PrPs_ratio_Tp2 == Makino_PrPs_ratio_Tp3)

print(are_equal) 
```