### 2025-05-07 CO2_response_fitting/Scripts/msuRACiFit_rice_params.Rmd
Update and rename 03_msuRACiFit_rice.Rmd to msuRACiFit_rice_params.Rmd

---
title: "msuRACiFits_fits_rice_params"
author: "Supreeta Vijayakumar"
date: "`r Sys.Date()`"
output: html_document
---

## Install A/Ci fitting package if necessary

msuRACiFit is a package of r scripts used to fit CO2 response curves developed by Thomas Sharkey's lab at Michigan State University.\
\
Source Code:\
<https://github.com/poales/msuRACiFit>

Publication:\
<https://doi.org/10.1111/pce.14153>

### Install the package using devtools:

```{r}
library(devtools)
devtools::install_github("poales/msuRACiFit")
```

### or remotes:

```{r}
remotes::install_github("poales/msuRACiFit")
```

### Load the here, msuRACiFit and readxl libraries:

```{r}
library(here)
library(readxl)
library(msuRACiFit)
```

Using \`here' roots the file path to the project environment.

### Import Excel files containing gas exchange data:

```{r}
A330502WT1 <- read_excel(here("Data","IR64-A009-07-33-05-02_Wildtype1.xlsx"))
A330502WT2 <- read_excel(here("Data","IR64-A009-07-33-05-02_Wildtype2.xlsx"))
A330504WT1 <- read_excel(here("Data","IR64-A009-07-33-05-04_Wildtype1.xlsx"))
A330504WT2 <- read_excel(here("Data","IR64-A009-07-33-05-04_Wildtype2.xlsx"))
A330506WT1 <- read_excel(here("Data","IR64-A009-07-33-05-06_Wildtype1.xlsx"))
A330506WT2 <- read_excel(here("Data","IR64-A009-07-33-05-06_Wildtype2.xlsx"))
A330509WT1 <- read_excel(here("Data","IR64-A009-07-33-05-09_Wildtype1.xlsx"))
A330509WT2 <- read_excel(here("Data","IR64-A009-07-33-05-09_Wildtype2.xlsx"))
```

### Store the files in a list:

```{r}
datasets <- list(c(A330502WT1),
                 c(A330502WT2),
                 c(A330504WT1),
                 c(A330504WT2),
                 c(A330506WT1),
                 c(A330506WT2),
                 c(A330509WT1),
                 c(A330509WT2))
```

### Calculate Gamma Star at different temperatures using calcGammaStar:

```{r}
msuRACiFit::calcGammaStar(c,dHa,tleaf,oxygen)
```

The tool uses preset in-vitro values from Hermida-Carrera et al (2016), where everything is presented in ppm, except for N. tabacum and A. thaliana.

Since we are analyzing the A/Ci in Pa, we have to adjust the resulting Gamma\*, Kc and Ko that we get in ppm to Pa before running calculations.

$c$ and $\Delta H_a$ for Gamma\* presented in the app/ preset table are for ppm, and gamma\* is corrected as follows:\
\
Gamma\* (Pa) = Gamma\*(ppm) \* 1000 / 1000000 \* Patm,\
Kc (Pa) = Kc(umol mol -1) \* 1000 / 1000000 \* Patm\
Kcair (Pa) = Kcair(umol mol -1) \* 1000 / 1000000 \* Patm,\
Ko (kPa) = O(Pa) / ((Kcair(Pa) / Kc(Pa)) - 1)/1000

### Create vectors of mean temperature and pressure:

```{r}
mean_temps <- numeric(length = 8) 
mean_press <- numeric(length = 8)
```

```{r}
for (i in seq_along(datasets)) { 
  mean_temps[i] <- mean(datasets[[i]]$Tleaf) 
  mean_press[i] <- mean(datasets[[i]]$Press) }
```

### Call calcgammastar function with different temperature inputs and get outputs saved in single vector for 8 files:

```{r}
Gstars_rice<-msuRACiFit::calcGammaStar(13.7,24.6,mean_temps,21) #rice c and dHa 
Gstars_rice_25<-msuRACiFit::calcGammaStar(13.7,24.6,25,21) #rice c and dHa 
```

### Next, we account for the variation in Press for each dataset and converting GammaStar back to Pa:

```{r}
Gstars_Pa_rice<-Gstars_rice*(1000 / 1000000)*mean_press 
```

### Save these gamma star values in a .csv:

```{r}
write.csv(Gstars_rice,file = "Outputs/Parameters/Gstars_rice.csv") #ppm
```

```{r}
write.csv(Gstars_Pa_rice,file = "Outputs/Parameters/Gstars_Pa_rice.csv") #Pa
```

### Alternatively, to fit data using the Shiny app, run:

```{r}
msuRACiFit::genApp()
```

Select *O.sativa* as preset plant and fill in the average Tleaf and Pa for each curve, generating guesses and fits, saving outputs as .png and csv tables.

#### Calculate Kc and Kcair for our Tleaf values in umol mol-1 then convert them into Pa before working out Ko.

Make sure all constants are named differently within the function to what they are called outside the function.

### Import data for Kc_c and Kc_dHa:

```{r}
Kc_c <- 38.9
Kc_dHa <- 83.1
```

### Import data for Kcair_c and Kcair_dHa:

```{r}
Kcair_c <- 30.5 
Kcair_dHa <- 60.5
```

### Write function to use the mean temperature in celsius from each curve for calculating Kc and Kcair:

```{r}
kinetic_params <- function(tempinC) 
    { ideal_gas <- 0.008314 # Ideal gas constant 
    kelvin <- 273.15 # Kelvin constant 
    Kc_scaling <- Kc_c # Kc scaling constant (unitless)
    Kc_activation <- Kc_dHa # Kc energy of activation (kJ mol -1) 
    Kcair_scaling <- Kcair_c # Kcair scaling constant (unitless) 
    Kcair_activation <- Kcair_dHa # Kcair energy of activation (kJ mol -1) 
    O2 <- 21000 # Oxygen concentration (Pa) 
    # Calculate Kc in umol at a specific temperature 
    Kc <- exp(Kc_scaling - (Kc_activation / (ideal_gas * (kelvin + tempinC)))) 
    # Calculate Kcair in umol at a specific temperature 
    Kcair <- exp(Kcair_scaling - (Kcair_activation / (ideal_gas * (kelvin + tempinC))))      
    return(c(Kc,Kcair)) }
```

### Create an empty dataframe to store all Kc and Kcair outputs:

```{r}
all_param <- data.frame(Kc = numeric(length(mean_temps)), 
                        Kcair = numeric(length(mean_temps)))
```

### Loop through kinetic_params function, calculate Kc and Ko values at each temperature in mean_temps and convert values from ppm to Pa according to each mean_press:

```{r}
for (j in seq_along(mean_temps)) { 
  # Call the kinetic_params function to calculate Kc and Kcair for the current temperature and convert them to Pa based on mean pressure 
  param <- kinetic_params(mean_temps[j])*(1000 / 1000000)*mean_press[j] 
  # Store the results in the output dataframe 
  all_param[j, ] <- param }
```

### Currently Kc and Kcair are in Pa

### Append a column for calculating Ko from Kc and Kcair and converting these values to kPa:

```{r}
all_param$Ko<-(21000 / ((all_param$Kcair / all_param$Kc) - 1))/1000 
all_param$Temperature <- mean_temps 
all_param$Press <- mean_press 
write.csv(all_param,
         file = "Outputs/Parameters/fitComplete/Rice/kinetic_parameters_rice.csv")
```

### Input Kc and Ko values from all_param into fitComplete

### Call fitComplete with different inputs and use a list of lists to store outputs

### Initialise an empty list to store output of fitComplete:

```{r}
complete_Fits_rice_list <- vector("list", length(datasets)) 
```

### This loop indexes through vectors (mean_temps,mean_press,Gstars_Pa_rice) and lists (assimilation_rates, intercellular_CO2) to fit 8 curves using fitComplete and save their outputs

### Vector of forceValues/bound_l/bound_h in order: VcMax, J, TPU, gm, Rd, aG, aS

```{r}
for (k in seq_along(datasets)) {  
  data_k <- datasets[[k]]
  if ("data" %in% colnames(data_k)) {
    data_k <- data_k$data  
  }
  data_k <- as.data.frame(data_k)

  complete_Fits_rice <- fitComplete(
    data = data_k,  
    name_assimilation = "Photo",
    name_ci = "Ci",
    gammastar = Gstars_Pa_rice[k],
    O2 = 21,
    pressure = mean_press[k],
    tleaf = mean_temps[k],
    initialGuess = NA,
    forceValues = c(NA, NA, NA, NA, NA, NA, NA),
    bound_l = c(1, 1, 1, .001, .001, 0, 0),
    bound_h = c(1000, 1000, 1000, 30, 10, 1, .75),
    ignoreTPU = F,
    maxiter = 500,
    Kc = all_param[k, "Kc"],
    Ko = all_param[k, "Ko"]
  )

  complete_Fits_rice_list[[k]] <- complete_Fits_rice
}
```

### Fit curves individually using fitComplete algorithm:

```{r}
completeFit1_rice <- fitComplete(A330502WT1, 
                                name_assimilation = "Photo", 
                                name_ci = "Ci", 
                                gammastar = Gstars_Pa_rice[1], 
                                O2 = 21, 
                                pressure = mean_press[1], 
                                tleaf = mean_temps[1], 
                                initialGuess = NA,
                                forceValues = c(NA, NA, NA, NA, NA, NA, NA), 
                                bound_l = c(1, 1, 1, .001, .001, 0, 0), 
                                bound_h = c(1000, 1000, 1000, 30, 10, 1, .75), 
                                ignoreTPU = F, 
                                maxiter = 500, 
                                Kc = all_param[1, "Kc"], 
                                Ko = all_param[1, "Ko"]) 
```

```{r}
completeFit2_rice <- fitComplete(A330502WT2, 
                                 name_assimilation = "Photo", 
                                 name_ci = "Ci", 
                                 gammastar = Gstars_Pa_rice[2], 
                                 O2 = 21, 
                                 pressure = mean_press[2], 
                                 tleaf = mean_temps[2], 
                                 initialGuess = NA, 
                                 forceValues = c(NA, NA, NA, NA, NA, NA, NA),
                                 bound_l = c(1, 1, 1, .001, .001, 0, 0), 
                                 bound_h = c(1000, 1000, 1000, 30, 10, 1, .75), 
                                 ignoreTPU = F, maxiter = 500, 
                                 Kc = all_param[2, "Kc"], 
                                 Ko = all_param[2, "Ko"]) 
```

```{r}
completeFit3_rice <- fitComplete(A330504WT1, 
                                 name_assimilation = "Photo", 
                                 name_ci = "Ci", 
                                 gammastar = Gstars_Pa_rice[3], 
                                 O2 = 21, 
                                 pressure = mean_press[3], 
                                 tleaf = mean_temps[3], 
                                 initialGuess = NA, 
                                 forceValues = c(NA, NA, NA, NA, NA, NA, NA), 
                                 bound_l = c(1, 1, 1, .001, .001, 0, 0), 
                                 bound_h = c(1000, 1000, 1000, 30, 10, 1, .75), 
                                 ignoreTPU = F, 
                                 maxiter = 500, 
                                 Kc = all_param[3, "Kc"], 
                                 Ko = all_param[3, "Ko"]) 
```

```{r}
completeFit4_rice <- fitComplete(A330504WT2, 
                                 name_assimilation = "Photo", 
                                 name_ci = "Ci", 
                                 gammastar = Gstars_Pa_rice[4], 
                                 O2 = 21, 
                                 pressure = mean_press[4], 
                                 tleaf = mean_temps[4], 
                                 initialGuess = NA, 
                                 forceValues = c(NA, NA, NA, NA, NA, NA, NA), 
                                 bound_l = c(1, 1, 1, .001, .001, 0, 0), 
                                 bound_h = c(1000, 1000, 1000, 30, 10, 1, .75), 
                                 ignoreTPU = F, 
                                 maxiter = 500, 
                                 Kc = all_param[4, "Kc"], 
                                 Ko = all_param[4, "Ko"]) 
```

```{r}
completeFit5_rice <- fitComplete(A330506WT1, 
                                 name_assimilation = "Photo", 
                                 name_ci = "Ci", 
                                 gammastar = Gstars_Pa_rice[5], 
                                 O2 = 21, 
                                 pressure = mean_press[5], 
                                 tleaf = mean_temps[5], 
                                 initialGuess = NA, 
                                 forceValues = c(NA, NA, NA, NA, NA, NA, NA), 
                                 bound_l = c(1, 1, 1, .001, .001, 0, 0), 
                                 bound_h = c(1000, 1000, 1000, 30, 10, 1, .75), 
                                 ignoreTPU = F, 
                                 maxiter = 500, 
                                 Kc = all_param[5, "Kc"], 
                                 Ko = all_param[5, "Ko"])
```

```{r}
completeFit6_rice <- fitComplete(A330506WT2, 
                                 name_assimilation = "Photo", 
                                 name_ci = "Ci", 
                                 gammastar = Gstars_Pa_rice[6], 
                                 O2 = 21, 
                                 pressure = mean_press[6], 
                                 tleaf = mean_temps[6], 
                                 initialGuess = NA, 
                                 forceValues = c(NA, NA, NA, NA, NA, NA, NA), 
                                 bound_l = c(1, 1, 1, .001, .001, 0, 0), 
                                 bound_h = c(1000, 1000, 1000, 30, 10, 1, .75), 
                                 ignoreTPU = F, 
                                 maxiter = 500, 
                                 Kc = all_param[6, "Kc"], 
                                 Ko = all_param[6, "Ko"]) 
```

```{r}
completeFit7_rice <- fitComplete(A330509WT1, 
                                 name_assimilation = "Photo", 
                                 name_ci = "Ci", 
                                 gammastar = Gstars_Pa_rice[7], 
                                 O2 = 21, 
                                 pressure = mean_press[7], 
                                 tleaf = mean_temps[7], 
                                 initialGuess = NA, 
                                 forceValues = c(NA, NA, NA, NA, NA, NA, NA), 
                                 bound_l = c(1, 1, 1, .001, .001, 0, 0), 
                                 bound_h = c(1000, 1000, 1000, 30, 10, 1, .75), 
                                 ignoreTPU = F, 
                                 maxiter = 500, 
                                 Kc = all_param[7, "Kc"], 
                                 Ko = all_param[7, "Ko"]) 
```

```{r}
completeFit8_rice <- fitComplete(A330509WT2, 
                                 name_assimilation = "Photo", 
                                 name_ci = "Ci", 
                                 gammastar = Gstars_Pa_rice[8], 
                                 O2 = 21, pressure = mean_press[8], 
                                 tleaf = mean_temps[8], 
                                 initialGuess = NA, 
                                 forceValues = c(NA, NA, NA, NA, NA, NA, NA), 
                                 bound_l = c(1, 1, 1, .001, .001, 0, 0), 
                                 bound_h = c(1000, 1000, 1000, 30, 10, 1, .75), 
                                 ignoreTPU = F, 
                                 maxiter = 500, 
                                 Kc = all_param[8, "Kc"], 
                                 Ko = all_param[8, "Ko"]) 
```

### Save parameter outputs:

```{r}
write.csv(completeFit1_rice[[1]], 
          file = "Outputs/Parameters/fitComplete/Rice/completefit1_console_rice_params.csv")

write.csv(completeFit2_rice[[1]], 
          file = "Outputs/Parameters/fitComplete/Rice/completefit2_console_rice_params.csv")

write.csv(completeFit3_rice[[1]], 
          file = "Outputs/Parameters/fitComplete/Rice/completefit3_console_rice_params.csv")

write.csv(completeFit4_rice[[1]], 
          file = "Outputs/Parameters/fitComplete/Rice/completefit4_console_rice_params.csv")

write.csv(completeFit5_rice[[1]], 
          file = "Outputs/Parameters/fitComplete/Rice/completefit5_console_rice_params.csv")

write.csv(completeFit6_rice[[1]], 
          file = "Outputs/Parameters/fitComplete/Rice/completefit6_console_rice_params.csv")

write.csv(completeFit7_rice[[1]], 
          file = "Outputs/Parameters/fitComplete/Rice/completefit7_console_rice_params.csv")

write.csv(completeFit8_rice[[1]], 
          file = "Outputs/Parameters/fitComplete/Rice/completefit8_console_rice_params.csv")
```

## Produce data tables for each curve:

```{r}
completeTables1_rice <- reconstituteTable(data=A330502WT1, 
                                          fitParams = completeFit1_rice[[1]],
                                          name_assimilation = "Photo", 
                                          name_ci=c("Ci"), 
                                          gammastar=Gstars_Pa_rice[1],
                                          O2=21, 
                                          pressure=mean_press[(1)], 
                                          tleaf = mean_temps[(1)], 
                                          ignoreTPU=F, 
                                          Kc=all_param[1, "Kc"], 
                                          Ko=all_param[1, "Ko"])

completeTables2_rice <- reconstituteTable(data=A330502WT2, 
                                          fitParams = completeFit2_rice[[1]], 
                                          name_assimilation = "Photo", 
                                          name_ci=c("Ci"), 
                                          gammastar=Gstars_Pa_rice[2], 
                                          O2=21, 
                                          pressure=mean_press[(2)], 
                                          tleaf = mean_temps[(2)], 
                                          ignoreTPU=F, 
                                          Kc=all_param[2, "Kc"],
                                          Ko=all_param[2, "Ko"])

completeTables3_rice <- reconstituteTable(data=A330504WT1, 
                                          fitParams = completeFit3_rice[[1]], 
                                          name_assimilation = "Photo", 
                                          name_ci=c("Ci"), 
                                          gammastar=Gstars_Pa_rice[3], 
                                          O2=21, 
                                          pressure=mean_press[(3)], 
                                          tleaf = mean_temps[(3)], 
                                          ignoreTPU=F, 
                                          Kc=all_param[3, "Kc"], 
                                          Ko=all_param[3, "Ko"])

completeTables4_rice <- reconstituteTable(data=A330504WT2, 
                                          fitParams = completeFit4_rice[[1]], 
                                          name_assimilation = "Photo",
                                          name_ci=c("Ci"), 
                                          gammastar=Gstars_Pa_rice[4], 
                                          O2=21, 
                                          pressure=mean_press[(4)], 
                                          tleaf = mean_temps[(4)], 
                                          ignoreTPU=F, 
                                          Kc=all_param[4, "Kc"], 
                                          Ko=all_param[4, "Ko"])

completeTables5_rice <- reconstituteTable(data=A330506WT1, 
                                          fitParams = completeFit5_rice[[1]], 
                                          name_assimilation = "Photo", 
                                          name_ci=c("Ci"), 
                                          gammastar=Gstars_Pa_rice[5], 
                                          O2=21, 
                                          pressure=mean_press[(5)], 
                                          tleaf = mean_temps[(5)], 
                                          ignoreTPU=F, 
                                          Kc=all_param[5, "Kc"], 
                                          Ko=all_param[5, "Ko"])

completeTables6_rice <- reconstituteTable(data=A330506WT2, 
                                          fitParams = completeFit6_rice[[1]], 
                                          name_assimilation = "Photo", 
                                          name_ci=c("Ci"), 
                                          gammastar=Gstars_Pa_rice[6], 
                                          O2=21, 
                                          pressure=mean_press[(6)], 
                                          tleaf = mean_temps[(6)], 
                                          ignoreTPU=F, 
                                          Kc=all_param[6, "Kc"], 
                                          Ko=all_param[6, "Ko"])

completeTables7_rice <- reconstituteTable(data=A330509WT1, 
                                          fitParams = completeFit7_rice[[1]], 
                                          name_assimilation = "Photo", 
                                          name_ci=c("Ci"), 
                                          gammastar=Gstars_Pa_rice[7], 
                                          O2=21, 
                                          pressure=mean_press[(7)], 
                                          tleaf = mean_temps[(7)], 
                                          ignoreTPU=F, 
                                          Kc=all_param[7, "Kc"], 
                                          Ko=all_param[7, "Ko"])

completeTables8_rice <- reconstituteTable(data=A330509WT2, 
                                          fitParams = completeFit8_rice[[1]], 
                                          name_assimilation = "Photo", 
                                          name_ci=c("Ci"), 
                                          gammastar=Gstars_Pa_rice[8], 
                                          O2=21, 
                                          pressure=mean_press[(8)], 
                                          tleaf = mean_temps[(8)], 
                                          ignoreTPU=F, 
                                          Kc=all_param[8, "Kc"], 
                                          Ko=all_param[8, "Ko"])
```

### Save data tables:

```{r}
write.csv(completeTables1_rice, 
          file = "Outputs/Tables/fitComplete/Rice/completefit1_console_rice_dataTable.csv") 
write.csv(completeTables2_rice, 
          file = "Outputs/Tables/fitComplete/Rice/completefit2_console_rice_dataTable.csv")
write.csv(completeTables3_rice, 
          file = "Outputs/Tables/fitComplete/Rice/completefit3_console_rice_dataTable.csv") 
write.csv(completeTables4_rice, 
          file = "Outputs/Tables/fitComplete/Rice/completefit4_console_rice_dataTable.csv")
write.csv(completeTables5_rice, 
          file = "Outputs/Tables/fitComplete/Rice/completefit5_console_rice_dataTable.csv")
write.csv(completeTables6_rice, 
          file = "Outputs/Tables/fitComplete/Rice/completefit6_console_rice_dataTable.csv") 
write.csv(completeTables7_rice, 
          file = "Outputs/Tables/fitComplete/Rice/completefit7_console_rice_dataTable.csv")
write.csv(completeTables8_rice, 
          file = "Outputs/Tables/fitComplete/Rice/completefit8_console_rice_dataTable.csv") 
```

### Plot graphs of limitations:

```{r}
completeGraph1_rice <- reconstituteGraph(data=A330502WT1, 
                                         fitParams = completeFit1_rice[[1]], 
                                         name_assimilation="Photo", 
                                         name_ci=("Ci"), 
                                         gammastar=Gstars_Pa_rice[1], 
                                         O2=21, 
                                         pressure=mean_press[(1)], 
                                         tleaf=mean_temps[(1)], 
                                         ignoreTPU=F, 
                                         Kc=all_param[1, "Kc"], 
                                         Ko=all_param[1, "Ko"])

completeGraph2_rice <- reconstituteGraph(data=A330502WT2, 
                                         fitParams = completeFit2_rice[[1]], 
                                         name_assimilation="Photo", name_ci=("Ci"), 
                                         gammastar=Gstars_Pa_rice[2], O2=21, 
                                         pressure=mean_press[(2)], tleaf=mean_temps[(2)], 
                                         ignoreTPU=F,
                                         Kc=all_param[2, "Kc"], 
                                         Ko=all_param[2, "Ko"])

completeGraph3_rice <- reconstituteGraph(data=A330504WT1, 
                                         fitParams = completeFit3_rice[[1]], 
                                         name_assimilation="Photo", 
                                         name_ci=("Ci"), 
                                         gammastar=Gstars_Pa_rice[3], 
                                         O2=21, 
                                         pressure=mean_press[(3)], 
                                         tleaf=mean_temps[(3)], 
                                         ignoreTPU=F, 
                                         Kc=all_param[3, "Kc"], 
                                         Ko=all_param[3, "Ko"])

completeGraph4_rice <- reconstituteGraph(data=A330504WT2, 
                                         fitParams = completeFit4_rice[[1]], 
                                         name_assimilation="Photo", 
                                         name_ci=("Ci"), 
                                         gammastar=Gstars_Pa_rice[4],
                                         O2=21, 
                                         pressure=mean_press[(4)], 
                                         tleaf=mean_temps[(4)], 
                                         ignoreTPU=F, 
                                         Kc=all_param[4, "Kc"], 
                                         Ko=all_param[4, "Ko"])

completeGraph5_rice <- reconstituteGraph(data=A330506WT1, 
                                         fitParams = completeFit5_rice[[1]], 
                                         name_assimilation="Photo", 
                                         name_ci=("Ci"), 
                                         gammastar=Gstars_Pa_rice[5],
                                         O2=21, 
                                         pressure=mean_press[(5)], 
                                         tleaf=mean_temps[(5)], 
                                         ignoreTPU=F, 
                                         Kc=all_param[5, "Kc"], 
                                         Ko=all_param[5, "Ko"])

completeGraph6_rice <- reconstituteGraph(data=A330506WT2, 
                                         fitParams = completeFit6_rice[[1]], 
                                         name_assimilation="Photo", 
                                         name_ci=("Ci"), 
                                         gammastar=Gstars_Pa_rice[6],
                                         O2=21, 
                                         pressure=mean_press[(6)], 
                                         tleaf=mean_temps[(6)], 
                                         ignoreTPU=F, 
                                         Kc=all_param[6, "Kc"], 
                                         Ko=all_param[6, "Ko"])

completeGraph7_rice <- reconstituteGraph(data=A330509WT1, 
                                         fitParams = completeFit7_rice[[1]], 
                                         name_assimilation="Photo", 
                                         name_ci=("Ci"), 
                                         gammastar=Gstars_Pa_rice[7],
                                         O2=21, 
                                         pressure=mean_press[(7)], 
                                         tleaf=mean_temps[(7)], 
                                         ignoreTPU=F, 
                                         Kc=all_param[7, "Kc"], 
                                         Ko=all_param[7, "Ko"])

completeGraph8_rice <- reconstituteGraph(data=A330509WT2, 
                                         fitParams = completeFit8_rice[[1]], 
                                         name_assimilation="Photo", 
                                         name_ci=("Ci"), 
                                         gammastar=Gstars_Pa_rice[8],
                                         O2=21,
                                         pressure=mean_press[(8)], 
                                         tleaf=mean_temps[(8)], 
                                         ignoreTPU=F, 
                                         Kc=all_param[8, "Kc"], 
                                         Ko=all_param[8, "Ko"]) 
```

### Save plots:

```{r}
png("Outputs/Graphs/fitComplete/Rice/completefits1_console_rice.png", width = 519, height = 457) 
plot(completeGraph1_rice) 
dev.off()

png("Outputs/Graphs/fitComplete/Rice/completefits2_console_rice.png", width = 519, height = 457) 
plot(completeGraph2_rice) 
dev.off()

png("Outputs/Graphs/fitComplete/Rice/completefits3_console_rice.png", width = 519, height = 457) 
plot(completeGraph3_rice) 
dev.off()

png("Outputs/Graphs/fitComplete/Rice/completefits4_console_rice.png", width = 519, height = 457) 
plot(completeGraph4_rice) 
dev.off()

png("Outputs/Graphs/fitComplete/Rice/completefits5_console_rice.png", width = 519, height = 457) 
plot(completeGraph5_rice) 
dev.off()

png("Outputs/Graphs/fitComplete/Rice/completefits6_console_rice.png", width = 519, height = 457) 
plot(completeGraph6_rice) 
dev.off()

png("Outputs/Graphs/fitComplete/Rice/completefits7_console_rice.png", width = 519, height = 457) 
plot(completeGraph7_rice) 
dev.off()

png("Outputs/Graphs/fitComplete/Rice/completefits8_console_rice.png", width = 519, height = 457) 
plot(completeGraph8_rice) 
dev.off()
```

#### Calculate sum of squared residuals:

```{r}
SSR1<-sum(completeTables1_rice$res^2)
SSR2<-sum(completeTables2_rice$res^2)
SSR3<-sum(completeTables3_rice$res^2)
SSR4<-sum(completeTables4_rice$res^2)
SSR5<-sum(completeTables5_rice$res^2)
SSR6<-sum(completeTables6_rice$res^2)
SSR7<-sum(completeTables7_rice$res^2)
SSR8<-sum(completeTables8_rice$res^2)
```

### Combine and save SSR values:

```{r}
SSR_values <- data.frame(SSR1,SSR2,SSR3,SSR4,SSR5,SSR6,SSR7,SSR8) 
write.csv(SSR_values, 
          file = "Outputs/Parameters/fitComplete/Rice/rice_console_SSR_values.csv", 
          row.names = FALSE)
```

### Export plot coordinates for limitations, example below:

```{r}
Ac<-completeGraph1_rice$plot_env$cdat 
Aj<-completeGraph1_rice$plot_env$jdat 
Ap<-completeGraph1_rice$plot_env$pdat 
limitations <- list(Ac,Aj,Ap)

write.csv(limitations,"Outputs/Tables/fitComplete/Rice/completeFit1_rice_limitations.csv")
```